fdcoexist

### 2018-12-14 meta_simulations.Rmd
Working variation in H over k

n_patches <- 25   # Number of patches
n_species <- 100  # Number of species
n_gen     <- 150  # Length of model run (generations)
initpop   <- 50   # initial population size
n_traits  <- 3    # number of traits
k         <- 1.15 # Maximum growth rate
env_width <- rep(10, length.out = n_patches)  # Envtal filter strength
var_width <- seq(10, 1, length.out = n_patches) # Varying envtal filter strength
total_runs = 100  # Number of set of traits to test

# Competition + Dispersion coefficients
B = 2e-4 # Scalar for intraspecific competition
A = 2e-4 # Alpha scalar
d = 0.05 # Dispersal percentage

@@

    apply(all_facets, 1, function(given_row) {
        multigen(traits = all_cor[[given_row[["cor_level"]]]],
                 trait_weights = trait_scenarios[[given_row[["scenario"]]]],
                 env = env,
                 time = n_gen,
                 species = n_species,
                 patches = n_patches,
                 composition = composition,
                 A = all_compet[[given_row[["compet_status"]]]][["A"]],
                 B = all_compet[[given_row[["compet_status"]]]][["B"]],
                 d = d,
                 k = given_k,
                 width = all_env[[given_row[["env_width"]]]]) %>%
            landscape_df(all_cor[[given_row[["cor_level"]]]],
                         trait_scenarios[[given_row[["scenario"]]]],
                         time = 150) %>%
            inner_join(all_cor[[given_row[["cor_level"]]]] %>%
                           as.data.frame() %>%
                           rownames_to_column("species"),
                       by = "species") %>%
            group_by(patch) %>%
            mutate(rel_abund = N/sum(N)) %>%
            ungroup() %>%
            mutate(compet_status = given_row[["compet_status"]],
                   env_width     = given_row[["env_width"]],
                   cor_level     = given_row[["cor_level"]],
                   scenario      = given_row[["scenario"]],
                   seed_number = seed_number) %>%
            select(seed_number, compet_status, env_width, cor_level, scenario,
                   everything())
    }) %>%
        bind_rows() %>%
        filter(N != 0) %>%  # Take out absent species
        mutate_at(vars(contains("trait")), funs(cwm = rel_abund * .)) %>%
        mutate(scenario = fct_relevel(scenario, names(trait_scenarios)),
               patch_optim = as.numeric(gsub("patches", "", patch))) %>%
        group_by(seed_number, compet_status, env_width, cor_level, scenario,
                 patch) %>%
        summarise(trait1_cwm  = sum(trait1_cwm),
                  trait2_cwm  = sum(trait2_cwm),
                  trait3_cwm  = sum(trait3_cwm),
                  trait1_var  = var(trait1),
                  trait2_var  = var(trait2),
                  trait3_var  = var(trait3),
                  patch_optim = unique(patch_optim))
}

@@

    all_simuls = apply(all_facets, 1, function(given_row) {
        multigen(traits = all_cor[[given_row[["cor_level"]]]],
                 trait_weights = trait_scenarios[[given_row[["scenario"]]]],
                 env = env,
                 time = n_gen,
                 species = n_species,
                 patches = n_patches,
                 composition = composition,
                 A = all_compet[[given_row[["compet_status"]]]][["A"]],
                 B = all_compet[[given_row[["compet_status"]]]][["B"]],
                 d = d,
                 k = given_k,
                 width = all_env[[given_row[["env_width"]]]])
    #     %>%
    #         landscape_df(all_cor[[given_row[["cor_level"]]]],
    #                      trait_scenarios[[given_row[["scenario"]]]],
    #                      time = 150) %>%
    #         inner_join(all_cor[[given_row[["cor_level"]]]] %>%
    #                        as.data.frame() %>%
    #                        rownames_to_column("species"),
    #                    by = "species") %>%
    #         group_by(patch) %>%
    #         mutate(rel_abund = N/sum(N)) %>%
    #         ungroup() %>%
    #         mutate(compet_status = given_row[["compet_status"]],
    #                env_width     = given_row[["env_width"]],
    #                cor_level     = given_row[["cor_level"]],
    #                scenario      = given_row[["scenario"]],
    #                seed_number = seed_number) %>%
    #         select(seed_number, compet_status, env_width, cor_level, scenario,
    #                everything())
    })
    # %>%
    #     bind_rows() %>%
    #     filter(N != 0) %>%  # Take out absent species
    #     mutate_at(vars(contains("trait")), funs(cwm = rel_abund * .)) %>%
    #     mutate(scenario = fct_relevel(scenario, names(trait_scenarios)),
    #            patch_optim = as.numeric(gsub("patches", "", patch))) %>%
    #     group_by(seed_number, compet_status, env_width, cor_level, scenario,
    #              patch) %>%
    #     summarise(trait1_cwm  = sum(trait1_cwm),
    #               trait2_cwm  = sum(trait2_cwm),
    #               trait3_cwm  = sum(trait3_cwm),
    #               trait1_var  = var(trait1),
    #               trait2_var  = var(trait2),
    #               trait3_var  = var(trait3),
    #               patch_optim = unique(patch_optim))

    list(simul = all_simuls, trait = uncor_traits)
}
# Generate traits --------------------------------------------------------------
trait_seed = 5
# Make traits with different correlations
set.seed(trait_seed)
uncor_traits = generate_cor_traits(n_patches, n_species, n_traits - 1,
                                   cor_coef = 0)
set.seed(trait_seed)
low_cor_traits = generate_cor_traits(n_patches, n_species, n_traits - 1,
                                     cor_coef = 0.3)

@@

    meta_simul(x, given_k = normal_k)
}, .progress = FALSE)
tictoc::toc()
```

@@

>>>>

n_patches <- 25   # Number of patches
n_species <- 100  # Number of species
n_gen     <- 150  # Length of model run (generations)
initpop   <- 50   # initial population size
n_traits  <- 3    # number of traits
k         <- 1.15 # Maximum growth rate
env_width <- rep(10, length.out = n_patches)  # Envtal filter strength
var_width <- seq(10, 1, length.out = n_patches) # Varying envtal filter strength
total_runs = 20  # Number of set of traits to test

# Competition + Dispersion coefficients
B = 2e-4       # Scalar for intraspecific competition
A = 2e-4       # Alpha scalar
d = 0.05       # Dispersal percentage
dist_power = 2 # Power of trait dissimilarity

# Hierarchical coefficients
H = 1.01
th_max = 25

@@

    apply(all_facets, 1, function(given_row) {
        multigen(traits = all_cor[[given_row[["cor_level"]]]],
                 trait_weights = trait_scenarios[[given_row[["scenario"]]]],
                 env = env,
                 time = n_gen,
                 species = n_species,
                 patches = n_patches,
                 composition = composition,
                 A = all_compet[[given_row[["compet_status"]]]][["A"]],
                 B = all_compet[[given_row[["compet_status"]]]][["B"]],
                 d = d,
                 k = given_k,
                 dist_power = dist_power,
                 H = H, th_max = th_max,
                 width = all_env[[given_row[["env_width"]]]]) %>%
            landscape_df(all_cor[[given_row[["cor_level"]]]],
                         trait_scenarios[[given_row[["scenario"]]]],
                         time = 150) %>%
            inner_join(all_cor[[given_row[["cor_level"]]]] %>%
                           as.data.frame() %>%
                           rownames_to_column("species"),
                       by = "species") %>%
            group_by(patch) %>%
            mutate(rel_abund = N/sum(N)) %>%
            ungroup() %>%
            mutate(compet_status = given_row[["compet_status"]],
                   env_width     = given_row[["env_width"]],
                   cor_level     = given_row[["cor_level"]],
                   scenario      = given_row[["scenario"]],
                   seed_number = seed_number) %>%
            select(seed_number, compet_status, env_width, cor_level, scenario,
                   everything())
    }) %>%
        bind_rows() %>%
        filter(N != 0) %>%  # Take out absent species
        mutate_at(vars(contains("trait")), funs(cwm = rel_abund * .)) %>%
        mutate(scenario = fct_relevel(scenario, names(trait_scenarios)),
               patch_optim = as.numeric(gsub("patches", "", patch))) %>%
        group_by(seed_number, compet_status, env_width, cor_level, scenario,
                 patch) %>%
        summarise(trait1_cwm  = sum(trait1_cwm),
                  trait2_cwm  = sum(trait2_cwm),
                  trait3_cwm  = sum(trait3_cwm),
                  trait1_var  = var(trait1),
                  trait2_var  = var(trait2),
                  trait3_var  = var(trait3),
                  patch_optim = unique(patch_optim))
}

@@

    all_simuls = apply(all_facets, 1, function(given_row) {
        multigen(traits = all_cor[[given_row[["cor_level"]]]],
                 trait_weights = trait_scenarios[[given_row[["scenario"]]]],
                 env = env,
                 time = n_gen,
                 species = n_species,
                 patches = n_patches,
                 composition = composition,
                 A = all_compet[[given_row[["compet_status"]]]][["A"]],
                 B = all_compet[[given_row[["compet_status"]]]][["B"]],
                 d = d,
                 k = given_k,
                 width = all_env[[given_row[["env_width"]]]])
    })

    list(simul = all_simuls, trait = uncor_traits)
}

meta_simul_A_B_k_and_H = function(seed_number, given_k = k, given_A = A,
                                  given_B_over_A = B/A,
                                  given_H_over_k = H/k) {
    set.seed(seed_number)
    uncor_traits = generate_cor_traits(n_patches, n_species, n_traits - 1,
                                       cor_coef = 0)
    all_cor = list(uncor = uncor_traits)

    all_compet = list(compet = list(A = given_A, B = given_A * given_B_over_A,
                                    H = given_k * given_H_over_k))

    our_scenars = list(
    R50A50   = data.frame(trait = paste0("trait", 1:3),
                          growth_weight = c(0.5, 0.5, 0),
                          compet_weight = c(0,   0.5, 0.5)))

    all_facets = expand.grid(compet_status = names(all_compet),
                             env_width     = names(all_env),
                             cor_level     = names(all_cor),
                             scenario      = names(our_scenars))

    apply(all_facets, 1, function(given_row) {
        multigen(traits = all_cor[[given_row[["cor_level"]]]],
                 trait_weights = trait_scenarios[[given_row[["scenario"]]]],
                 env = env,
                 time = n_gen,
                 species = n_species,
                 patches = n_patches,
                 composition = composition,
                 A = all_compet[[given_row[["compet_status"]]]][["A"]],
                 B = all_compet[[given_row[["compet_status"]]]][["B"]],
                 d = d,
                 k = given_k,
                 dist_power = dist_power,
                 H = all_compet[[given_row[["compet_status"]]]][["H"]],
                 th_max = th_max,
                 width = all_env[[given_row[["env_width"]]]]) %>%
            landscape_df(all_cor[[given_row[["cor_level"]]]],
                         trait_scenarios[[given_row[["scenario"]]]],
                         time = 150) %>%
            inner_join(all_cor[[given_row[["cor_level"]]]] %>%
                           as.data.frame() %>%
                           rownames_to_column("species"),
                       by = "species") %>%
            group_by(patch) %>%
            mutate(rel_abund = N/sum(N)) %>%
            ungroup() %>%
            mutate(compet_status = given_row[["compet_status"]],
                   env_width     = given_row[["env_width"]],
                   cor_level     = given_row[["cor_level"]],
                   scenario      = given_row[["scenario"]],
                   seed_number = seed_number) %>%
            select(seed_number, compet_status, env_width, cor_level, scenario,
                   everything())
    }) %>%
        bind_rows() %>%
        mutate_at(vars(contains("trait")), funs(cwm = rel_abund * .)) %>%
        mutate(scenario = fct_relevel(scenario, names(trait_scenarios)),
               patch_optim = as.numeric(gsub("patches", "", patch))) %>%
        group_by(seed_number, compet_status, env_width, cor_level, scenario,
                 patch) %>%
        summarise(species_rich = sum(N > 0),
                  trait1_cwm   = sum(trait1_cwm),
                  trait2_cwm   = sum(trait2_cwm),
                  trait3_cwm   = sum(trait3_cwm),
                  trait1_var   = var(trait1),
                  trait2_var   = var(trait2),
                  trait3_var   = var(trait3),
                  patch_optim  = unique(patch_optim))
}
# Generate traits --------------------------------------------------------------
trait_seed = 5
# Make traits with different correlations
set.seed(trait_seed)
uncor_traits = generate_cor_traits(n_patches, n_species, n_traits - 1,
                                   cor_coef = 0)
set.seed(trait_seed)
low_cor_traits = generate_cor_traits(n_patches, n_species, n_traits - 1,
                                     cor_coef = 0.3)

@@

    meta_simul(x, given_k = normal_k)
}, .progress = TRUE)
tictoc::toc()
```

@@

## Coexistence plot depending on A, B/A and k

```{r param_space_simuls}
list_A = 10^-(seq(2, 6, length.out = 5))
list_k = seq(1, 1.5, length.out = 5)
list_B_over_A = 10^seq(0, 4, length.out = 5)
list_H_over_k = seq(0, 1, length.out = 5)

var_a_var_k_var_ba = cross(list(run_n = seq(1), A = list_A, k = list_k,
                         B_over_A = list_B_over_A,
                         H_over_k = list_H_over_k)) %>%
    future_map(function(x) {
    devtools::load_all()

    meta_simul_A_B_k_and_H(x$run_n, x$k, x$A, x$B_over_A,
                           given_H_over_k = x$H_over_k) %>%
        mutate(k = x$k, A = x$A, B_over = x$B_over_A, H_over_k = x$H_over_k)
}, .progress = TRUE)
tictoc::toc()

var_ba_df = bind_rows(var_a_var_k_var_ba)
```

```{r interactive_species_rich_plot}
library("plotly")

var_ba_df %>%
    filter(patch == "patches13", species_rich > 0) %>%
    plot_ly(x = ~k, y = ~A, z = ~B_over, color = ~species_rich,
            text = ~paste("SR: ", species_rich)) %>%
    add_markers() %>%
    layout(scene = list(xaxis = list(title = ),
                        yaxis = list(title = 'A (interspecific competition)',
                                     type = "log"),
                        zaxis = list(title = 'B/A (ratio intra/inter)',
                                     type = "log")))
```

```{r biplot_of_parameters_richness}
plot_k_A = var_ba_df %>%
    filter(patch %in% paste0("patches", c(1, 13))) %>%
    ggplot(aes(k, A)) +
    stat_summary_2d(aes(z = species_rich), fun = "mean", geom = "raster") +
    facet_wrap(~patch) +
    scale_fill_viridis_c() +
    scale_y_log10() +
    labs(x = "k (maximum growth rate)",
         y = "A (interspecific competition)",
         fill = "Species Richness\n(at t = 150)",
         subtitle = "Average Richness over all values of B/A & H/k") +
    theme(legend.position = "top")

plot_A_BA = var_ba_df %>%
    filter(patch %in% paste0("patches", c(1, 13))) %>%
    ggplot(aes(A, B_over)) +
    stat_summary_2d(aes(z = species_rich), fun = "mean", geom = "raster") +
    facet_wrap(~patch) +
    scale_fill_viridis_c() +
    scale_x_log10() +
    scale_y_log10() +
    labs(x = "A (interspecific competition)",
         y = "B/A (ratio intra/inter)",
         fill = "Species Richness\n(at t = 150)",
         subtitle = "Average Richness over all values of k & H/k") +
    theme(legend.position = "top")

plot_k_BA = var_ba_df %>%
    filter(patch %in% paste0("patches", c(1, 13))) %>%
    ggplot(aes(k, B_over)) +
    stat_summary_2d(aes(z = species_rich), fun = "mean", geom = "raster") +
    facet_wrap(~patch) +
    scale_fill_viridis_c() +
    scale_y_log10() +
    labs(x = "k (maximum growth rate)",
         y = "B/A (ratio intra/inter)",
         fill = "Species Richness\n(at t = 150)",
         subtitle = "Average Richness over all values of A & H/k") +
    theme(legend.position = "top")

plot_k_Hk = var_ba_df %>%
    filter(patch %in% paste0("patches", c(1, 13))) %>%
    ggplot(aes(k, H_over_k)) +
    stat_summary_2d(aes(z = species_rich), fun = "mean", geom = "raster") +
    facet_wrap(~patch) +
    scale_fill_viridis_c() +
    labs(x = "k (maximum growth rate)",
         y = "H/k (ratio hierarchical compet/max. growth)",
         fill = "Species Richness\n(at t = 150)",
         subtitle = "Average Richness over all values of A and B/A") +
    theme(legend.position = "top")

plot_A_Hk = var_ba_df %>%
    filter(patch %in% paste0("patches", c(1, 13))) %>%
    ggplot(aes(A, H_over_k)) +
    stat_summary_2d(aes(z = species_rich), fun = "mean", geom = "raster") +
    facet_wrap(~patch) +
    scale_fill_viridis_c() +
    scale_x_log10() +
    labs(x = "A (interspecific competition)",
         y = "H/k (ratio hierarchical compet/max. growth)",
         fill = "Species Richness\n(at t = 150)",
         subtitle = "Average Richness over all values of k and B/A") +
    theme(legend.position = "top")

plot_BA_Hk = var_ba_df %>%
    filter(patch %in% paste0("patches", c(1, 13))) %>%
    ggplot(aes(B_over, H_over_k)) +
    stat_summary_2d(aes(z = species_rich), fun = "mean", geom = "raster") +
    facet_wrap(~patch) +
    scale_fill_viridis_c() +
    scale_x_log10() +
    labs(x = "B/A (ratio intra/inter)",
         y = "H/k (ratio hierarchical compet/max. growth)",
         fill = "Species Richness\n(at t = 150)",
         subtitle = "Average Richness over all values of k and A") +
    theme(legend.position = "top")

plot_grid(plot_k_A,
          plot_k_Hk,
          plot_A_BA,
          plot_A_Hk,
          plot_k_BA,
          plot_BA_Hk,
          ncol = 2)
```

```{r biplot_with_R2}
var_ba_r2 = var_ba_df %>%
    ungroup() %>%
    filter(!is.nan(trait2_cwm)) %>%
    nest(-k, -A,-B_over, -H_over_k, -seed_number, -scenario) %>%
    mutate(data = data %>%
               map(~.x %>%
                       gather("cwm_name", "cwm_value", ends_with("cwm")) %>%
                       group_by(cwm_name, add = TRUE) %>%
                       do(mod_cwm_env = lm(cwm_value ~ patch_optim, data = .)) %>%
                       ungroup() %>%
                       mutate(mod_glance = lapply(mod_cwm_env, broom::glance)) %>%
                       unnest(mod_glance))) %>%
    unnest(data) %>%
    filter(cwm_name == "trait2_cwm")

var_labs = c(k        = "k (maximum growth rate)",
             A        = "A (interspecific competition)",
             B_over   = "B/A (ratio intra/inter)",
             H_over_k = "H/k (ratio hierarch./max. growth)")

# function to plot R² of CWM <-> Env relationship
plot_average_r_squared = function(var_df, x_var, y_var, given_labs = var_labs) {

    not_used_vars = setdiff(names(given_labs), c(x_var, y_var))

    var_df %>%
        ggplot(aes_string(x_var, y_var)) +
    stat_summary_2d(aes(z = adj.r.squared), fun = "mean", geom = "raster") +
    scale_fill_viridis_c() +
    labs(x = var_labs[[x_var]],
         y = var_labs[[y_var]],
         fill = "CWM <-> Env. R^2",
         subtitle = paste0("Average R-squared over all values of ",
                           paste(not_used_vars, collapse = " and "))) +
    theme(legend.position = "top",
          legend.text = element_text(angle = 45))
}

# Plotting R-squared
plot_k_A_r2 = plot_average_r_squared(var_ba_r2, "k", "A") +
    scale_y_log10()

plot_A_BA_r2 = plot_average_r_squared(var_ba_r2, "A", "B_over") +
    scale_x_log10() +
    scale_y_log10()

plot_k_BA_r2 = plot_average_r_squared(var_ba_r2, "k", "B_over") +
    scale_y_log10()

plot_k_Hk_r2 = plot_average_r_squared(var_ba_r2, "k", "H_over_k")

plot_A_Hk_r2 = plot_average_r_squared(var_ba_r2, "A", "H_over_k") +
    scale_x_log10()

plot_BA_Hk_r2 = plot_average_r_squared(var_ba_r2, "B_over", "H_over_k") +
    scale_x_log10()

plot_grid(plot_k_A_r2,
          plot_k_Hk_r2,
          plot_A_BA_r2,
          plot_A_Hk_r2,
          plot_k_BA_r2,
          plot_BA_Hk_r2,
          ncol = 2)
```

```{r biplot_with_slopes}
var_ba_slope = var_ba_df %>%
    ungroup() %>%
    filter(!is.nan(trait2_cwm)) %>%
    nest(-k, -A,-B_over, -H_over_k, -seed_number, -scenario) %>%
    mutate(data = data %>%
               map(~.x %>%
                       gather("cwm_name", "cwm_value", ends_with("cwm")) %>%
                       group_by(cwm_name, add = TRUE) %>%
                       do(mod_cwm_env = lm(cwm_value ~ patch_optim, data = .)) %>%
                       ungroup() %>%
                       mutate(mod_glance = lapply(mod_cwm_env, broom::tidy)) %>%
                       unnest(mod_glance))) %>%
    unnest(data) %>%
    filter(cwm_name == "trait2_cwm") %>%
    filter(term == "patch_optim")

plot_average_slope = function(var_df, x_var, y_var, given_labs = var_labs) {

    not_used_vars = setdiff(names(given_labs), c(x_var, y_var))

    var_df %>%
        ggplot(aes_string(x_var, y_var)) +
        stat_summary_2d(aes(z = estimate), fun = "mean", geom = "raster") +
        scale_fill_viridis_c() +
        labs(x = var_labs[[x_var]],
             y = var_labs[[y_var]],
             fill = "CWM <-> Env. Slope",
             subtitle = paste0("Average R-squared over all values of ",
                               paste(not_used_vars, collapse = " and "))) +
        theme(legend.position = "top",
          legend.text = element_text(angle = 45))
}

# Plotting CWM <-> Env. Slope
plot_k_A_slope = plot_average_slope(var_ba_slope, "k", "A") +
    scale_y_log10()

plot_A_BA_slope = plot_average_slope(var_ba_slope, "A", "B_over") +
    scale_x_log10() +
    scale_y_log10()

plot_k_BA_slope = plot_average_slope(var_ba_slope, "k", "B_over") +
    scale_y_log10()

plot_k_Hk_slope = plot_average_slope(var_ba_slope, "k", "H_over_k")

plot_A_Hk_slope = plot_average_slope(var_ba_slope, "A", "H_over_k") +
    scale_x_log10()

plot_BA_Hk_slope = plot_average_slope(var_ba_slope, "B_over", "H_over_k") +
    scale_x_log10()

plot_grid(plot_k_A_slope,
          plot_k_Hk_slope,
          plot_A_BA_slope,
          plot_A_Hk_slope,
          plot_k_BA_slope,
          plot_BA_Hk_slope,
          nrow = 3)
```
